Towards Patient-Specific Neural Field Modeling using Micro-Electrode Data

### 2011-02-07 ltx/data_paper/DataPaper.tex
edits to M step

Note that $\boldsymbol\Sigma_e=\sigma_d^2\tilde{\boldsymbol\Sigma}_e$ where $\tilde{\boldsymbol\Sigma}_e$ is constant and can be calculated beforehand as the support of the disturbance covariance function is assumed to be known. Substituting equations \ref{eq:CondititionDist1} and \ref{eq:CondititionDist2} into \ref{eq:QIntermsofJointDist} and taking trace and rearranging, using the invariant cyclic permutations property of the trace, this distribution can be written as
\begin{align}\label{eq:Qfunctiontrace}

@@

 2 \ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta ,\sigma_d) &=-\mathbf x_{t+1}^\top\boldsymbol\Sigma_e^{-1}\mathbf x_{t+1}+2\mathbf x_{t+1}^\top\boldsymbol\Sigma_e^{-1}q( \mathbf x_t)\boldsymbol\theta+2\xi \mathrm{tr} \left\lbrace \mathbf x_t\mathbf x_{t+1}^\top\boldsymbol\Sigma_e^{-1}\right\rbrace \nonumber \\
&-\boldsymbol\theta^\top q^\top(\mathbf x_t)\boldsymbol\Sigma_e^{-1}q(\mathbf x_t)\boldsymbol\theta-2\xi \mathbf x_t^\top\boldsymbol\Sigma_e^{-1}q(\mathbf x_t)\boldsymbol\theta-\xi^2\mathrm{tr}\left\lbrace \mathbf x_t \mathbf x_t^\top\boldsymbol\Sigma_e^{-1}\right\rbrace. 

@@

Having approximated $q\left(\mathbf x_t\right)$ in the E-step, Substituting \ref{eq:qTaylor} in \ref{eq:Qfunctionintrace} and taking the expectation of the log likelihood function for all time instants gives an approximate to the lower-bound that needs to be maximised to gain optimal parameter estimates. Therefore $\mathcal Q$-function can be approximated as  

@@

 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&\approx& -(T-1)\ln \mid\boldsymbol\Sigma_{\epsilon}\mid-\mathrm{tr}\left\lbrace\boldsymbol\Sigma_{\epsilon}^{-1}\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace\nonumber \\
&+&\beta+2\Xi_0\boldsymbol\theta+
2\xi \mathrm{tr}\left\lbrace \Xi_1 \boldsymbol\Sigma_e^{-1}\right\rbrace \nonumber \\
&-&\boldsymbol\theta^\top \Xi_2\boldsymbol\theta-2\xi \Xi_3 \boldsymbol\theta - \xi^2 \mathrm{tr} \left\lbrace\Xi_4\boldsymbol\Sigma_e^{-1} \right\rbrace 

@@

	\beta &= \sum_{t=0}^{T-1}\mathbf E_{\Theta'}\left[-\mathbf x_{t+1}^\top\boldsymbol\Sigma_e^{-1}\mathbf x_{t+1}\right] \label{eq:def of beta} \\
\Xi_0&=\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t+1}^\top\boldsymbol\Sigma_e^{-1}q(\mathbf{\hat x}_t)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf M_{t+1} \boldsymbol\Sigma_e^{-1}  \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t) d\mathbf{r}'\right] \\	

@@

\Xi_2&=\sum_{t=0}^{T-1}\left[q^\top(\mathbf{\hat x}_t)\boldsymbol\Sigma_e^{-1}q(\mathbf{\hat x}_t)+\Lambda_t(\mathbf r)\mathbf P_t \tilde{\Lambda}_t(\mathbf r)\right] \\
 \Xi_3&=\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t}^\top\boldsymbol\Sigma_e^{-1}q(\mathbf{\hat x}_t)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf P_t \boldsymbol\Sigma_e^{-1}  \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t) d\mathbf{r}'\right] \\

@@

	 \boldsymbol\lambda_t^{(i)}(\mathbf r)&=\int_{\Omega} \boldsymbol \Psi^\top(\mathbf r')\boldsymbol \Sigma_e^{-1}\phi_i(\mathbf r')f'(\boldsymbol\phi^\top(\mathbf r')\mathbf {\hat x}_t)d\mathbf r'\\

>>>>

Note that $\boldsymbol\Sigma_e=\sigma_d^2\tilde{\boldsymbol\Sigma}_e$ where $\tilde{\boldsymbol\Sigma}_e$ is constant and can be calculated beforehand as the support of the disturbance covariance function is assumed to be known. Taking trace and rearranging, using the invariant cyclic permutations property of the trace, this distribution can be written as
\begin{align}\label{eq:Qfunctiontrace1}

@@

 2 \ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta ,\sigma_d) &=-n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace\mathbf x_{t+1} \mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace+\frac{2}{\sigma_d^2}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\boldsymbol\theta+\frac{2\xi}{\sigma_s^2} \mathrm{tr} \left\lbrace \mathbf x_t\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace \nonumber \\
&-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top q^\top(\mathbf x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\boldsymbol\theta-\frac{2\xi}{\sigma_d^2} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\boldsymbol\theta-\frac{\xi^2}{\sigma_d^2}\mathrm{tr}\left\lbrace \mathbf x_t \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace. \label{eq:Qfunctiontrace2}

@@

Note that we ignore $\ln\mid\tilde{\boldsymbol\Sigma_e}\mid $ as it is not a function of the parameter set. Having approximated $q\left(\mathbf x_t\right)$ in the E-step, by substituting  \ref{eq:Qfunctiontrace1} and \ref{eq:Qfunctiontrace2} into \ref{eq:QIntermsofJointDist} and taking the expectation of the log likelihood function for all time instants gives an approximate to the lower-bound that needs to be maximised to gain optimal parameter estimates. Therefore $\mathcal Q$-function can be approximated as  

@@

 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&\approx& -(T-1)n_y\ln \sigma_{\epsilon}^2-\mathrm{tr}\left\lbrace\boldsymbol\Sigma_{\epsilon}^{-1}\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace\nonumber \\
&-&(T-1)n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace \Xi'_0 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +2\frac{1}{\sigma_d^2}\Xi_0\boldsymbol\theta+
\frac{2\xi}{\sigma_d^2} \mathrm{tr}\left\lbrace \Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace \nonumber \\
&-&\frac{1}{\sigma_d^2}\boldsymbol\theta^\top \Xi_2\boldsymbol\theta-\frac{2\xi}{\sigma_d^2} \Xi_3 \boldsymbol\theta -  \frac{\xi^2}{\sigma_d^2}\mathrm{tr} \left\lbrace\Xi_4\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace 

@@

	\Xi'_0&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}\mathbf x_{t+1}^\top\right]=\sum_{t=0}^{T-1}\left(\mathbf P_{t+1}+\mathbf{\hat x}_{t+1}\mathbf{\hat x}_{t+1}^\top\right)\label{eq:def of Xi'0} \\
\Xi_0&=\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf M_{t+1} \tilde{\boldsymbol\Sigma}_e^{-1} \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t) d\mathbf{r}'\right] \\	

@@

\Xi_2&=\sum_{t=0}^{T-1}\left[q^\top(\mathbf{\hat x}_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\Lambda_t(\mathbf r)\mathbf P_t \tilde{\Lambda}_t(\mathbf r)\right] \\
 \Xi_3&=\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf P_t \tilde{\boldsymbol\Sigma}_e^{-1}  \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t) d\mathbf{r}'\right] \\

@@

	 \boldsymbol\lambda_t^{(i)}(\mathbf r)&=\int_{\Omega} \boldsymbol \Psi^\top(\mathbf r')\tilde{\boldsymbol\Sigma}_e^{-1}\phi_i(\mathbf r')f'(\boldsymbol\phi^\top(\mathbf r')\mathbf {\hat x}_t)d\mathbf r'\\
